### 2023-10-31 ATACCC-appendix-original-analysis.tex
Bulk of ATACCC chapter (no results yet)

\documentclass[thesis.tex]{subfiles}

\begin{document}

\chapter{Original ATACCC analysis} \label{ATACCC:sec:original-analysis}

\end{document}