### 2023-12-14 backcalc.tex
Add backcalc chapter

\documentclass[thesis.tex]{subfiles}

\title{Backcalculation}
\author{Joshua Blake}
\date{\today}

\begin{document}

\ifSubfilesClassLoaded{
  \setcounter{chapter}{6}
}

\chapter{Estimating incidence using backcalculation} \label{backcalc}

Now that I have estimated the duration of PCR positivity, I can estimate the incidence of SARS-CoV-2.
In this chapter, I will apply the theory introduced in \cref{E-incidence-prev} to the CIS data.
I use the period from the expansion of the CIS, starting 31st August 2020, until there was significant vaccination rollout\todo{insert date}.
First, I explain the method I will apply (\cref{})

\section{Methods}

\section{Results}

\section{Discussion}

\section{Conclusion}

\ifSubfilesClassLoaded{
  \listoftodos
}{}

\end{document}