### 2023-12-20 backcalc.tex

\documentclass[thesis.tex]{subfiles}

\title{Backcalculation}
\author{Joshua Blake}
\date{\today}

\begin{document}

\ifSubfilesClassLoaded{
  \setcounter{chapter}{6}
}

\chapter{Estimating incidence using backcalculation} \label{backcalc}

\todo[inline]{
    There is an asymmetry with this chapter and the following one on SEIR modelling.
    In this chapter, the theory/background is explained in an introductory/background chapter, with a short chapter with the application-specifics and results here.
    In the following chapter, the theory/background is explained within the chapter which means it is much longer.
    I'm not sure if this is an issue? But this chapter seems quite short currently and the next one quite long.
    I could move the SEIR theory to the background chapter, and possibly combine this chapter with the next one into a single chapter on incidence and transmission estimation?
    This would make sense for allowing a comparison between the results into its own section as well, rather than currently it being placed slightly awkwardly in the results and discussion sections of the SEIR chapter.
}

\section{Conclusion} \label{backcalc:sec:conclustion}

\ifSubfilesClassLoaded{
  \listoftodos
}{}

\end{document}