### 2024-02-14 transmission-appendix-INLA.tex
add INLA appendix

\documentclass[thesis.tex]{subfiles}

\begin{document}

\end{document}