### 2024-02-19 transmission-appendix-INLA.tex
start writing INLA appendix

\documentclass[thesis.tex]{subfiles}

\begin{document}

\chapter{INLA} \label{transmission:sec:INLA}

INLA (Integrated Nested Laplace Approximation) is a method for approximate Bayesian inference.
It was introduced by \textcite{rueINLA}.
This appendix is based on that paper, as well as the seminar \textcite{rueINLAseminar} and the tutorial paper \textcite{martinoINLAtutorial}.

\section{Latent Gaussian model}

The INLA method requires a LGM (latent Gaussian model).
An LGM is defined in terms of a vector of hyperparameters, $\vec{\varphi}$; a deterministic transformation of the hyperparameters into covariance matrix, $\matr{\Sigma}(\vec{\varphi})$; a vector of latent variables, $\vec{x}$; a vector of $n$ data points, $\vec{y} = [y_1, \dots, y_n]^T$; and $n$ linear predictors $\vec{\eta} = [\eta_1, \dots, \eta_n]^T$.
To be an LGM, $\vec{x}$ must be a latent Gaussian field, that is $\vec{x} \mid \vec{\varphi} \dist \N(\vec{0}, \matr{\Sigma}(\vec{\varphi}))$; and the likelihood must have the conditional independence property $p(\vec{y} \mid \vec{\varphi}, \vec{x}) = \prod_j p(y_j \mid \eta_j(\vec{x}), \vec{\varphi})$.

The Gaussian structure allows the use of the approximations in \cref{transmission:sec:INLA:inference}.
The conditional independence structure means that the precision matrix, $\matr{Q} = \matr{\Sigma}^{-1}$, is sparse.
The sparsity of $\matr{Q}$ allows for efficient computation of the procedure.

\section{Inference procedure} \label{transmission:sec:INLA:inference}

\subsection{Gaussian likelihood}

Start by the case in which the likelihood, $\vec{y} \mid \vec{x}, \vec{\theta}$, is Gaussian.
In this case, as $\vec{x} \mid \vec{\theta}$ is also Gaussian, the distributions of $\vec{x}, \vec{y} \mid \vec{\theta}$ and $\vec{x} \mid \vec{y}, \vec{\theta}$ are both Gaussian too.

The posterior density of $\vec{\theta}$, $\vec{\theta }\mid \vec{y}$, is not Gaussian.
Its density (up to a normalising constant) is:
\begin{align}
p(\vec{\theta} \mid \vec{y}) \propto \frac{p(\vec{x}, \vec{y} \mid \vec{\theta}) p(\vec{\theta})}{p(\vec{x} \mid \vec{y}, \vec{\theta})}.
\end{align}
All of the densities on the right-hand side can easily be found analytically.

Assume that $\vec{\theta} \mid \vec{y}$ is unimodal, which is very likely for reasonable priors on $\vec{\theta}$.
Find its mode via optimisation and explore its area, until the density is negligible.
By integrating over this area, the normalising constant can be found.
Hence, the posterior density of the hyperparameters is known.

Finally, the marginal posterior density of $\vec{x}$ is found by integrating over $\vec{\theta}$:
\begin{align}
p(\vec{x} \mid \vec{y}) = \int p(\vec{x} \mid \vec{\theta}) p(\vec{\theta} \mid \vec{y}) d\vec{\theta}.
\end{align}
which can be computed using a Monte Carlo integration over $p(\vec{\theta} \mid \vec{y})$, previously calculated.

\subsection{Non-Gaussian likelihood}

Now, the INLA approach needs extending to non-Gaussian, but known, likelihoods for $\vec{y} \mid \vec{x}, \vec{\theta}$.
The density $p(\vec{x}, \vec{y} \mid \vec{\theta})$ is not Gaussian, but still known.
However, $\vec{x} \mid \vec{y}, \vec{\theta}$ is not Gaussian, and its density is not known.

INLA assumes that $\vec{x} \mid \vec{y}, \vec{\theta}$ is approximately Gaussian, and therefore takes a Laplace approximation to the distribution.
The greater the smoothing, the better the approximation.
This is because, if $\vec{x} \mid \vec{y}, \vec{\theta}$ is then dominated by the prior $\vec{x} \mid \vec{\theta}$ and the prior is Gaussian.
\todo[inline]{Add justification for the Laplace approximation}

\section{Drawing from the posterior} \label{transmission:sec:INLA:posterior}

\section{Default priors} \label{transmission:sec:INLA:priors}

\end{document}