### 2024-03-14 conclusion.tex
first draft of conclusion

\documentclass[thesis.tex]{subfiles}

\begin{document}
\ifSubfilesClassLoaded{
    \setcounter{chapter}{7}
}

\chapter{Conclusions} \label{conclusion}

Pandemics will continue to pose a significant threat to public health for the foreseeable future.
Therefore, there is a public health need to respond to pandemics.
Such a response needs to be informed by monitoring.
In this thesis, I have investigated one method of monitoring pandemics: prevalence surveys.

\section{Main findings}

In \cref{E-intro:sec:aims}, I posed the central question for this thesis: can prevalence surveys alone be used to infer incidence and transmission, and if so how?

In \cref{E-transmission}, I found that this it is possible, developing two models that can infer incidence and transmission from prevalence surveys.
One model was statistical in nature and the other mechanistic.
The statistical approach, based on the backcalculation framework, estimates incidence and makes fewer assumptions about the infection process; however, the estimates are very uncertain.
The mechanistic approach, within the compartmental modelling framework, reduces the uncertainty and allows estimation of the components of transmission by making stronger assumptions about the infection process.
These results highlighted the heterogeneity across regions and age groups in England.

The work in \cref{E-transmission} relied on the theoretical work of \cref{E-inc-prev}.
This work developed the backcalculation framework for the context of prevalence surveys for SARS-CoV-2.
The major insight was the justification of the deterministic model framework and the importance of the distribution of the duration of positivity.

These models both required knowledge of the distribution of the duration of positivity.
Estimating this distribution was the focus of the majority of this thesis.
In \cref{E-perf-test,E-imperf-test}, I showed how this quantity could be estimated the CIS prevalence survey.
This did require external information from the ATACCC study (the results of \cref{E-ATACCC}), demonstrating that the CIS's design is not sufficient for estimating transmission.
However, future surveys may be able to avoid this issue (see \cref{conclusion:sec:future-work}).

Prior to this thesis, the duration of positivity in the general population was not well understood (see \cref{E-intro:sec:previous-duration-estimates}).
A particularly understudied area was quantifying the shape and size of the tail of the distribution, due to limited follow-up of individuals who test positive.

Overall, I have found that representative prevalence surveys are a promising approach for pandemic surveillance.
A well-designed survey is sufficient to estimate all of the quantities of interest (as described in \cref{E-intro:sec:metrics}), except severity.

\section{Future work} \label{conclusion:sec:future-work}

The feasibility of jointly modelling duration and incidence was not investigated in this thesis.
The model for duration in \cref{E-perf-test,E-imperf-test} estimated the total number of infections in the cohort but assumed these occurred uniformly over the duration of the survey.
If, rather than assuming uniformity, the time of incidence was estimated, this would be a form of incidence estimation.
The issue of including covariates to control for non-response bias, encountered in \cref{E-transmission}, as well as the parameterisation of the shape of the incidence curve would be major challenges in this extension.
The basis of such a model could be existing modelling frameworks~\autocite[e.g.][]{taffeJoint,haySerosolver}.
These frameworks addressed the challenges by: taking approximations (\eg empirical Bayes), having smaller sample sizes, or using high-performance computing.
However, the CIS is larger and incidence estimation desired at a more granular level than previously encountered.
Furthermore, high-performance computing cannot be used within the SRS.

A common theme in this thesis was computational challenges.
The challenges could be addressed by more efficient MCMC algorithms, approximate methods, or alternate computing methods (\eg the use of GPUs).
Alternate computing methods would require appropriate hardware within the SRS, which should be considered when choosing the design of future TREs.

The CIS was a household-based survey, but the household structure was not considered in this thesis.
The household-based design is likely to induce clustering in results.
Therefore, the uncertainty in my results might be underestimated.
While the beta-binomial likelihood used in \cref{E-transmission} should capture this, it does not consider the clustering between different strata.
Other chapters did not consider clustering at all.
Similarly, the treatment of a longitudinal survey as a cross-sectional survey in \cref{E-perf-test} may have underestimated the uncertainty in the results.
More sophisticated incorporation of clustering, for instance using random effects for households and/or individuals, could be considered in future work.

The challenges in estimating the distribution of the duration of positivity originated in the design of the studies used.
Exploring the design of future studies to better estimate the distribution of the duration of positivity would be beneficial.
A particularly promising direction would be intensive follow-up of individuals following a positive test (\eg daily testing) until they have two or more negative tests.
Intensive follow-up would increase the precision by which the end of the infection episode is known.
Furthermore, it reduces the impact of false negatives because they are likely to be followed by a true positive.
I have developed software that could be used to perform simulation studies of alternative designs.

Other design questions are also posed by the work in this thesis.
For example, whether a representative sample of the population is optimal, or if certain groups should be oversampled.
Certain groups, such as younger individuals, were more likely to be infected.
It is plausible that sampling in proportion to incidence would be beneficial.
This is especially true if policy questions (\eg closing of schools) are of importance.
These questions could be investigated through simulation studies.
In constructing such studies, it will be important to consider how these considerations might be different for different pathogens (\eg influenza, the most common source of respiratory pandemics) and tests (\eg rapid antigen tests, which have a shorter duration of positivity).

This thesis did not consider detection or early surveillance of emerging pandemics.
Initial uncertainty about local transmission in the UK may have led to suboptimal decisions at the start of 2020~\autocite{pellisChallenges}.
Community testing was not widely available at this stage of the pandemic, limiting the knowledge of the extent of local (as opposed to imported) transmission~\autocite{whittyTechReportCOVID}.
A representative survey of the population could have provided this information.
However, challenges in constructing a survey with the power to detect low levels of transmission, as well as the time required to set-up such a survey, would need to be addressed.

A final consideration is the estimation of the severity of a pathogen.
Severity of a pathogen is crucial for risk assessment, and can change over time~\autocite[e.g.][]{kirwanSeverityTrends}.
Two approaches could be taken to using a prevalence surveys for this purpose.
First, a cohort approach where the outcomes of individuals who test positive are followed up.
The low number of positive cases in the CIS would make this challenging.
In addition, the data would need to be linked to relevant records.
Second, a transmission model, such as the one in \cref{E-SEIR}, could jointly model severe events and prevalence.
Such an approach has previously been used~\autocite{daviesAssociation,ironsEstimating,knockKey,nicholsonImproving,pooleyEstimation,birrellRTM2} but without the sophistication of the observation model developed in \cref{E-SEIR:sec:observation}.

\ifSubfilesClassLoaded{
  \listoftodos
}{}

\end{document}