### 2024-03-15 transmission-appendix-INLA.tex
finish info on INLA

\documentclass[thesis.tex]{subfiles}

\begin{document}

\chapter{INLA} \label{transmission:sec:INLA}

INLA (Integrated Nested Laplace Approximation) is a method for approximate Bayesian inference.
It was introduced by \textcite{rueINLA}.
This appendix is based on that paper, as well as the seminar \textcite{rueINLAseminar} and the tutorial paper \textcite{martinoINLAtutorial}.

The standard implementation of INLA is the R-INLA package~\autocite{RINLA}.

\section{Latent Gaussian model}

INLA requires a LGM (latent Gaussian model).
An LGM is defined in terms of a vector of hyperparameters, $\vec{\varphi}$; a deterministic transformation of the hyperparameters into covariance matrix, $\matr{\Sigma}(\vec{\varphi})$; a vector of latent variables, $\vec{x}$; a vector of $n$ data points, $\vec{y} = [y_1, \dots, y_n]^T$; and $n$ linear predictors $\vec{\eta} = [\eta_1, \dots, \eta_n]^T$.
To be an LGM, $\vec{x}$ must be a latent Gaussian field, that is $\vec{x} \mid \vec{\varphi} \dist \N(\vec{0}, \matr{\Sigma}(\vec{\varphi}))$; and the likelihood must have the conditional independence property $\pi(\vec{y} \mid \vec{\varphi}, \vec{x}) = \prod_j \pi(y_j \mid \eta_j(\vec{x}), \vec{\varphi})$.
% The Gaussian structure allows the use of the approximations in \cref{transmission:sec:INLA:inference}.
% The conditional independence structure means that the precision matrix, $\matr{Q} = \matr{\Sigma}^{-1}$, is sparse.

For INLA to perform well, \ie be compuutationally efficient and an accurate approximation, $\vec{\theta}$ should be of low dimension, \textcite{martinoINLAtutorial} recommend $< 15$; and the precision matrix, $\matr{Q}$, should be sparse.
The low dimensionality of $\vec{\theta}$ is because the method requires numerical integration over it.
The sparsity of $\matr{Q}$ allows for efficient computation.
Sparsity is introduced by conditional independence structures of the latent field that are present in many models.
For example, in the classic random effect setting, all random effects are independent conditional on the hyperparameters.

\section{Inference procedure} \label{transmission:sec:INLA:inference}

The core of INLA is an approximation for the posterior distribution of the hyperparameters $\pi(\vec{\theta} \mid \vec{y})$ and the marginal posterior distribution of the latent field, $\pi(x_i \mid \vec{y})$.

\subsection{Gaussian likelihood}

Start by the case in which the likelihood, $\vec{y} \mid \vec{x}, \vec{\theta}$, is Gaussian.
In this case, as $\vec{x} \mid \vec{\theta}$ is also Gaussian, the distributions of $\vec{x} \mid \vec{y}, \vec{\theta}$ is Gaussian too.
The Gaussian form of $\vec{x} \mid \vec{y}, \vec{\theta}$ follows from the conjugate property of the Gaussian distribution.
This distribution can easily be found analytically.

The posterior density of $\vec{\theta}$, $\vec{\theta }\mid \vec{y}$, is not Gaussian.
Approximate it by the distribution $\tilde\pi(\vec{\theta} \mid \vec{y})$ defined as:
\begin{align}
\tilde\pi(\vec{\theta} \mid \vec{y})
&\propto \frac{\pi(\vec{x^*}, \vec{y} \mid \vec{\theta}) \pi(\vec{\theta})}{\pi(\vec{x^*} \mid \vec{y}, \vec{\theta})}
\end{align}
where $\vec{x^*} = \argmax_{\vec{x}} \pi(\vec{x} \mid \vec{y}, \vec{\theta})$.
All of the densities on the right-hand side can easily be found analytically.

Assume that $\vec{\theta} \mid \vec{y}$ is unimodal, which is very likely for reasonable priors on $\vec{\theta}$.
Find its mode via optimisation and explore its area, until the density is negligible.
By integrating over this area, the normalising constant can be found.
Hence, the approximate posterior density of the hyperparameters is known.

Finally, the marginal posterior density of $x_i$ is found by integrating over $\vec{\theta}$:
\begin{align}
\pi(x_i \mid \vec{y})
&\approx \int \pi(x_i \mid \vec{\theta}, \vec{y}) \tilde\pi(\vec{\theta} \mid \vec{y}) d\vec{\theta}.
\end{align}
which can be computed using numerical integration over $\tilde\pi(\vec{\theta} \mid \vec{y})$, previously calculated.

\subsection{Non-Gaussian likelihood}

Now, the INLA needs extending to non-Gaussian, but known, likelihoods for $\vec{y} \mid \vec{x}, \vec{\theta}$.
The density $\pi(\vec{x}, \vec{y} \mid \vec{\theta})$ is not Gaussian, but still known.
However, $\vec{x} \mid \vec{y}, \vec{\theta}$ is not Gaussian, and its density is not known.

INLA assumes that $\vec{x} \mid \vec{y}, \vec{\theta}$ is approximately Gaussian, and therefore takes a Laplace approximation to the distribution.
The greater the smoothing, the better the approximation.
This is because, if $\vec{x} \mid \vec{y}, \vec{\theta}$ is then dominated by the prior $\vec{x} \mid \vec{\theta}$ and the prior is Gaussian.
Exploiting the structure of the problem, in particular the conditional independence structure, the Laplace approximation can be computed efficiently.

\section{Drawing from the posterior} \label{transmission:sec:INLA:posterior}

Various methods exist to sample from the joint posterior of the hyperparameters and the latent variables, based on the INLA approximation.
The joint posterior is not approximated by default.
I used the one implemented in the R-INLA function \texttt{inla.posterior.sample}, described in \textcite[section 4]{chiuchioloJoint}.

This method incorporates the first three moments of each marginal distribution, and uses a skew Gaussian copula to approximate the joint distribution.

\end{document}